Global code motion of congruent instructions

### 2017-11-08 cgo/paper.tex
move legality above

When SSAPRE inserts $\Phi$ (expression PHIs), it does so at places where
(potentially) multiple expressions may merge. If a $\Phi$ has a missing entry,
that means the expression is partially available.  We apply a similar concept
for hoisting expression by working on inverted graph. We insert a CHI, to a
basic block with multiple successors. Then for each instruction, we register its
flow of value to its corresponding CHI which can be found in its PDF. For
conciseness we only register values with multiple occurrences. Finally, we start
walking the CFG and remove the CHI if it has a missing entry, that means the
expression is not anticipable in the basic block having CHI. The completeness of
CHI is a necessary condition for hoistability but not sufficient. This is
because congruence of value numbers only implies that they compute same value if
the inputs are same. So for instructions like memory references, they may
compute different values should the memory get modified along the path. For that
we need to check for intersecting side-effects along the path.

@@

@@

\subsection{Legality check}
\label{subsec:legality}
Since the equality of candidates is purely based on the value numbers, we also
need to establish if hoisting them or sinking them would be legal. Once a common
dominator (post-dominator) is found, we check whether all the use-operands of
the set of instructions are available at that position. It is possible to
reinstantiate or remateralize the use-operands in some cases when the operands
are not available and make it legal to move the instruction.

\subsubsection{Downward-safety of movable instructions}

In order to be able to move an instruction that has side effects, we must
guarantee that the side-effects of that instruction before and after its
code-motion will produce the same side-effects.  For hoisting in particular, we
need to prove that on all execution paths from the hoisting point to the end of
the program the side effects appear in the same order. In general, we cannot
introduce a new computation along any path of execution without checking for
undefined behavior \cite{undef}. Since \GCM{} is only dealing with VBEs, as
hoisting happens for CHIs with complete entries, the downward safety is ensured
by construction of CHIs.

>>>>

When SSAPRE inserts $\Phi$ (expression PHIs), it does so at places where
(potentially) multiple expressions may merge. If a $\Phi$ has a missing entry,
that means the expression is partially available.  We apply a similar concept
for hoisting expression by working on inverted graph. We insert a CHI, to a
basic block with multiple successors. Then for each instruction, we register its
flow of value to its corresponding CHI which can be found in its PDF. For
conciseness we only register values with multiple occurrences. Finally, we start
walking the CFG and remove the CHI if it has a missing entry, that means the
expression is not anticipable in the basic block having CHI. For the remaining
CHIs anticipabilty of `values' is ensured by construction.

@@

\subsubsection{Legality of movable instructions}
\label{subsec:legality}
Since the equality of candidates is purely based on the value numbers, we also
need to establish if hoisting them or sinking them would be legal. The
completeness of CHI is a necessary condition for hoistability but not
sufficient. This is because congruence of value numbers only implies that they
compute same value if the inputs are same. So for instructions like memory
references, they may compute different values should the memory get modified
along the path. For that we need to check for intersecting side-effects along
the path.

For hoisting, we need to prove that on all execution paths from the hoisting
point to the end of the program the side effects appear in the same order. In
general, we cannot introduce a new computation along any path of execution
without checking for undefined behavior \cite{undef}. For that we need to see a)
if I aliases with any instruction before it in its basic block and b) Walk the
CFG from its basic block 'upwards' to its PDF and look for any interfering
side-effects. If there aren't any then it is safe to hoist.

@@
